---
title: "Tips and Tricks"
author: "Emmanel Pascual"
date: "2023-02-09"
categories: [Data, Tips, Tricks]
image: "cheatCode.avif"
---

### Here Comes The Great Tips and Tricks!

Look Up Functions\

```{r}
"?Class = This allows you to look up a function you have unpackaged"

"??Class = This allows you to look up a function you have not yet unpackaged"
```

Mean and Counting Tricks

1:5 is the same as 1,2,3,4,5 rep(x , y) x = Value , Y = instances , Rep = Repetition

```{r}
Yay <- rep(4 , 6)
Yay
vector1 <- c(1 , 2, 3 , 4, 5, 6 )
vector2 <- c((8:10))
vector1
vector2 
```

c() = combines values into a vector or a list (VECTOR) C( 1 , 2 , 3 )

```{r}
My_number <- c(1, 92, 182)
My_number
"We can Change the Vector using this My_number[2] <- 'two' "
My_number[2] <- "two"
"Lets see, background realness darling" 
My_number
"Remember when we add quotation marks to any Vectors the entire range becomes Strings & or Characters"

```

as.numeric() , as.character()

```{r}
"Since R is an interesting language we learned that it does not typecast nor conatonate" 
as.numeric(My_number)
"See if we have the string two instead of 2 an interger it doesn't work lets edit that out! "
My_number[2] <- 2 
as.numeric(My_number)
"Huzzah we did it! "
```

Mean(x, trim , na.rm) Trim

```{r}
numbs <- 1:6
mean(numbs)
"Trim = Data cleaning process rTrim() is very confusing. What the hell."
mean(numbs , .1 )

```

Sample function is random selection Sample (x , size , replace = FALSE Prob = Rep (⅙ , 6) Everything Sample makes a selection it wants to know what the probability is thats when the REP function comes into play.

```{r}
dice <- c(1:6)
dice
sample(dice, 6, replace = FALSE, prob = rep(1/6, 6))
```
